Mental Health of U.S. Citizens During the COVID-19 Pandemic

### 2022-03-21 outputs/paper/paper.Rmd
citations added

Trends were noticed in all segregations, with younger and poorer (by income) respondents showing significantly worse mental health. A comparison to respondents’ physical health was also made, with a notable number of respondents being physically healthy but mentally unhealthy.  Members of the LGBTQ+ community also show significantly poorer mental health than those of the heterosexual community. Finally, the social class was also assessed, based on self-identification; which yielded a similar trend to the income class (poorer had more mentally unhealthy respondents), except for the upper-most class; which showed a dramatic increase in poor mental health. These trends are crucial for the government, explicitly showcasing who needs the most help. With the help of policies, one could decrease the poor mental health rate in many of these classes, via analyzing which section of the class needs it most.

@@

To get better insights on what factors affect Mental health for different parts of the population, we utilized the 2021 US General Social Survey [@norc] from NORC at the University of Chicago [@norc-at-the-university-of-chicago]. The raw dataset is a Stata file and was imported in R using the package Haven. It includes data from 568 Survey Questions for 4032 Survey Participants. Using the R [@citeR] package tidyverse [@citetidyverse] and dplyr [@dplyr], we were able to clean and perform exploratory data analysis on the dataset to get insights into the data. Further, we used R package ggplot2 [@citeggplot2] to visualize the data for this paper.

@@

## Variables of interest

This survey contains a total of 565 variables and 4032 observations. However, due to the design of the survey and the impact of the Covid-19 pandemic, each participant may not receive the same survey questions thus leaving some responses empty.

>>>>

Trends were noticed in all segregations, with younger and poorer (by income) respondents showing significantly worse mental health. Men were also observed to have poorer mental health. A comparison to respondents’ physical health was also made, with a notable number of respondents being physically healthy but mentally unhealthy.  Members of the LGBTQ+ community also show significantly poorer mental health than those of the heterosexual community. Finally, the social class was also assessed, based on self-identification; which yielded a similar trend to the income class (poorer had more mentally unhealthy respondents), except for the upper-most class; which showed a dramatic increase in poor mental health. These trends are crucial for the government, explicitly showcasing who needs the most help. With the help of policies, one could decrease the poor mental health rate in many of these classes, via analyzing which section of the class needs it most.

@@

To get better insights on what factors affect Mental health for different parts of the population, we utilized the 2021 US General Social Survey [@norc] from NORC at the University of Chicago [@norc-at-the-university-of-chicago]. The raw dataset is a Stata file and was imported in R using the package Haven [@haven]. It includes data from 568 Survey Questions for 4032 Survey Participants. Using the R [@citeR] package tidyverse [@citetidyverse] and dplyr [@dplyr], we were able to clean and perform exploratory data analysis on the dataset to get insights into the data. Further, we used R package ggplot2 [@citeggplot2] to visualize the data for this paper.

@@

## Variables of Interest

This survey contains a total of 565 variables and 4032 observations. However, due to the design of the survey and the impact of the COVID-19 pandemic, each participant may not receive the same survey questions thus leaving some responses empty.
